Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility

### 2025-11-30 paper/04-analysis.qmd
Up antecedentes

## Descriptive statistics

@@

@fig-interact displays the origin-specific ATT estimates from the meritocracy interaction models as marginal effects of the mobility treatment at low and high meritocratic beliefs. Across trajectories, mobility effects for low-merit respondents are uniformly small and imprecise, hovering around zero for both upward and downward moves. Among high-merit respondents, the pattern aligns only partially with the proposed mechanism: upward mobility from middle to high status shows a clear and statistically significant pro-commodification effect ($\beta$ = 0.22, 95% CI [0.05, 0.38]), whereas all other trajectories—Low→Middle, Low→High, Middle→Low, High→Middle, and High→Low—remain statistically indistinguishable from zero or display negative estimates that do not systematically weaken under high merit. Formal interaction tests also fail to detect strong heterogeneity by meritocratic beliefs (see [Supplementary Material](#anexo) for interaction models). Taken together, the results provide only limited support for the mechanism hypothesis: meritocracy appears to amplify the pro-market consequences of upward mobility in one specific trajectory, but there is no general or consistent evidence that meritocratic beliefs constitute a stable channel through which intergenerational mobility shapes preferences for pension commodification.

>>>>

## Descriptive analysis

@@

@fig-interact displays the origin-specific ATT estimates from the meritocracy interaction models as marginal effects of the mobility treatment at low and high meritocratic beliefs. Across trajectories, mobility effects among low-merit respondents are uniformly small, imprecise, and centered around zero for both upward and downward moves. Among high-merit respondents, only upward mobility from middle to high status shows a clear pro-commodification pattern, with a positive and statistically significant effect on MJP for pensions ($\beta = 0.22$, 95% CI [0.05, 0.38]). All other trajectories—Low→Middle, Low→High, Middle→Low, High→Middle, and High→Low—remain statistically indistinguishable from zero. Nonetheless, the point estimates broadly align with the theoretical expectations for upward moves: upwardly mobile individuals with stronger meritocratic beliefs tend to display higher support for pension commodification than their low-merit counterparts. By contrast, downward trajectories exhibit a pattern opposite to the mechanism hypothesis, with high-merit respondents generally showing more de-commodifying attitudes than those with weaker meritocratic beliefs. However, formal interaction tests do not detect robust heterogeneity in mobility effects by meritocratic beliefs (see [Supplementary Material](#anexo) for full interaction models): even for the Middle→High trajectory, the simple effect under high meritocracy does not differ significantly from its low-merit counterpart. Taken together, these results do not provide convincing support for the mechanism hypothesis: meritocratic beliefs do not operate as a consistent or general channel through which intergenerational mobility shapes preferences for the commodification of pensions.
